Supporting Information: Population Effects of Marine Protected Areas Are Likely To Be Unclear

### 2020-04-22 documents/ovando-regional-effects-of-mpas-si-v2.Rmd
pre changing to population language commit

---
title: "Supporting Information: The Regional Effects of Marine Protected Areas"
author:  Daniel Ovando,Jennifer E. Caselle,Christopher Costello,Olivier Deschenes,Steven D. Gaines,Ray Hilborn,Owen Liu
output: 
  bookdown::pdf_document2:
    keep_tex: true
  pdf_document: default
bibliography: My Library.bib
header-includes:
    - \usepackage{longtable}
    - \usepackage{booktabs}
    - \usepackage{eso-pic}
    - \usepackage{graphicx}
    - \usepackage[left]{lineno}
    - \usepackage{xcolor}
    - \linenumbers
params:
  run_name: ["v6.0"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
always_allow_html: true
---

@@

All code needed to reproduce our main results and manuscript can be found at https://github.com/DanOvando/regional-effects-of-mpas. A fully reproducible environment for running this analysis and compiling the manuscript will be made available through Code Ocean at https://codeocean.com/capsule/5233105/tree. All analysis were performed in `r R.version.string`. Package versions are shown in Table.S\@ref(tab:pkg-info).

```{r pkg-info, results = "asis", eval=FALSE}
table <- devtools::package_info() %>% 
  select(package, loadedversion, date,source)

@@

PISCO data report positive observations of fish. In order to use these data in our model we need to add in zeros for any transect that could have observed a given species of fish but did not. We assume that a fish could have been observed on a given transect if that species has ever been observed at that site in any time period in the data (PISCO data are organized by sites, with multiple transect at different locations within the borders of a site). If a species has never been observed at a site we assume that it does not occur at that site.

@@

For each species at each transect, we then calculate the biomass density of that transect as the sum of the observed biomass divided by the transect area. We then average the biomass densities for each species across all the transects at a given site, and lastly sum these mean species biomass densities to achieve the total mean species biomass at the site level.

@@

To provide greater detail on this process, we first conduct the data pre-processing described earlier. From there, we aggregated data to the level of total biomass density of targeted and non-targeted species at each site each year. As such, our base model estiamtes the effect of the MPAs on total biomass density of targeted species, though we also explore the effect on alternative specifications such as mean biomass of targeted species. the model was then fit using the `rstanarm` package in R (5000 iterations, 2500 warmup, 4 chains, adapt_delta = 0.85).

The intercept prior was was determined using the `rstanarm` `autoscale` function. For the non-intercept terms, we manually set a normal(0,2) prior. This implies that coefficeints such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4. Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior.

The full table of covariates and their posterior means and 89% credible interval can be found in Table.S\@ref(tab:did-coefs).

```{r did-coefs, results = "asis",eval = FALSE}

did_fit %>% 
  broom::tidy(intervals = TRUE, prob = 0.89) %>% 
  select(term, estimate, lower, upper) %>% 
  rename(mean = estimate, lower_89th_percentile = lower, upper_89th_percentile= upper) %>% 
  knitr::kable(
      digits = 2,
            caption = "Posterior means and 89% credible interval for key model coefficients",
             longtable = TRUE,
             booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))
```

@@

ppc_dens_overlay(y = y,
                 yrep = yrep) +
  scale_x_continuous(name = "Log(Total Biomass Density)") +
  scale_colour_discrete(labels = c("Observed", "Posterior Predictive"),
                       name = '')

@@

```{r mean_test_plot, fig.cap = "Posterior predictive mean (distribution) and empirical mean of total biomass density. Posterior predictive mean adequately recovers empirical mean."}

@@

```{r sum_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately recovers empirical sum."}

@@

```{r sd_test_plot, fig.cap = "Posterior predictive standard deviation (distribution) and empirical standard deviation of total biomass density. Posterior predictive standard deviation adequately recovers empirical standard deviation"}
ppc_stat(y,  yrep, stat = "sd")

@@

```{r, fig.cap="Model fit to PISCO data but only using variables also available for KFM data", results="hide"}

@@

```{r, fig.cap="Model fit to KFM data. Results are more uncertain than using PISCO data, but follow the same trend estimated from the PISCO data."}

@@

@@

```{r, fig.cap="Estimated MPA effects using PISCO data from outside MPAs only. The base model uses data from both inside and outside MPAs. These results show that the same general trend holds when only using data from outside the MPAs, though the model estimates less probability of high positive effects witout MPA data, compared to the model fit using all the data.",results = "hide"}

@@

@@

@@

```{r mean-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish. For this run, we instead estimate the effect on mean biomass densities of targeted species. Results show the same pattern as our base run. "}
mean_fit <- model_runs %>%
  slice(1) %>%
  left_join(abundance_data, by = "data_source") %>%
  mutate(
    did_fit = pmap(
      list(
        data = data,
        data_to_use = data_to_use,
        data_source = data_source
      ),
      estimate_did,
      fit_mean = TRUE,
      iter = 2500,
      site_data = site_data,
      cdfw_catches = cdfw_catches,
      life_history_data = life_history_data,
      refresh = 1000
    )
  )

@@

```{r no-zero-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish. For this run, we instead estimate the effect on mean biomass densities of targeted species. Results show the same pattern as our base run. "}

@@

```
 selection on observables

@@

```{r betatable}

@@

Simulation results are presented as percent differences in biomass densities with and without MPAs, in order to be comparable to the estimates that the regression model produces. However, this metric presents some problems as a measure of how "detectable" an effect size is. As depletion increases, relatively small changes in total biomass (relative to the variance in the observation process) can translate into large percent changes in biomass. For example, moving from a density of .2kg/m^2^ to .4kg/m^2^ translates to a 100% percent increase, but only a .2kg/m^2^ absolute increase, a small value to detect with a real observation program.

>>>>

---
title: "Supporting Information: Population Effects of Marine Protected Areas Are Likely To Be Unclear"
author: 
  - Daniel Ovando
  - Jennifer E. Caselle
  - Christopher Costello
  - Olivier Deschenes
  - Steven D. Gaines
  - Ray Hilborn
  - Owen Liu
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2:
    keep_tex: true
  pdf_document: default
bibliography: My Library.bib
header-includes:
    - \usepackage{longtable}
    - \usepackage{booktabs}
    - \usepackage{eso-pic}
    - \usepackage{graphicx}
    - \usepackage[left]{lineno}
    - \usepackage{xcolor}
params:
  run_name: ["v6.0"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
always_allow_html: true
---

@@

All code needed to reproduce our main results and manuscript can be found at https://github.com/DanOvando/regional-effects-of-mpas. All analysis were performed in `r R.version.string`. Package versions are shown in Table.S\@ref(tab:pkg-info).

```{r pkg-info, results = "asis", eval=TRUE}
table <- devtools::package_info() %>% 
  select(package, loadedversion, date,source)

@@

PISCO data report positive observations of fish, in the manner of number of individuals of a species seen within a particular size group. In order to use these data in our model we need to add in zeros for any transect that could have observed a given species of fish but did not. We assume that a fish could have been observed on a given transect if that species has ever been observed at that site in any time period in the data (PISCO data are organized by sites, with multiple transect at different locations within the borders of a site). If a species has never been observed at a site we assume that it does not occur at that site.

@@

For each species at each transect, we calculate the biomass density of that transect as the sum of the observed biomass divided by the transect area. We then average the biomass densities for each species across all the transects at a given site, and lastly sum these mean species biomass densities to achieve the total mean species biomass at the site level.

@@

The full list of included coefficients and their estimated values can be seen in Table.\@ref(tab:did-coefs).

To provide greater detail on this process, we first conduct the data pre-processing described earlier. From there, we aggregated data to the level of total mean biomass density of targeted and non-targeted species at each site each year. As such, our base model estiamtes the effect of the MPAs on total mean biomass density of targeted species, though we also explore the effect on alternative specifications such as mean biomass of targeted species. the model was then fit using the `rstanarm` package in R (5000 iterations, 2500 warmup, 4 chains, adapt_delta = 0.85).

The intercept prior was was determined using the `rstanarm` `autoscale` function. For the non-intercept terms, we manually set a normal(0,2) prior. This implies that coefficeints such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4. Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior (e.g. a one standard deviation change in an independent variable changes the dependent variable by 4 on the log scale).

The full table of covariates and their posterior means and 89% credible interval can be found in Table.S\@ref(tab:did-coefs). "var" refers to variable, "_2" indicates a squared term. tex is the mean cumulative number of observations of the observers collecting the raw data, surge is the mean reported surge across transects, kelp is the mean kelp coverage at the site, temp is the water temperature, regional_temp_dev is the water temperature at a given site scaled by the mean water temperature at that site. Targeted is a boolean marker indicating whether the given observation is for targeted or non-targeted species.

```{r did-coefs, results = "asis",eval = TRUE}
did_fit %>%
  broom::tidy(intervals = TRUE, prob = 0.89) %>%
  select(term, estimate, lower, upper) %>%
  rename(
    mean = estimate,
    lower_89th_ci = lower,
    upper_89th_ci = upper
  ) %>%
  knitr::kable(
    digits = 2,
    caption = "Posterior means and credible interval for key model coefficients",
    longtable = TRUE,
    booktabs = TRUE ) %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "repeat_header"))

@@

ppc_dens_overlay(y = y,
                 yrep = yrep) +
  scale_x_continuous(name = "Total Biomass Density") +
  scale_colour_discrete(labels = c("Observed", "Posterior Predictive"),
                       name = '')

@@

```{r mean_test_plot, fig.cap = "Posterior predictive mean (distribution) and empirical mean of total biomass density. Posterior predictive mean adequately covers empirical mean."}

@@

```{r sum_test_plot, fig.cap = "Posterior predictive sum (distribution) and empirical sum of total biomass density. Posterior predictive sum adequately covers empirical sum."}

@@

```{r sd_test_plot, fig.cap = "Posterior predictive standard deviation (distribution) and empirical standard deviation of total biomass density. Posterior predictive standard deviation adequately covers empirical standard deviation"}
ppc_stat(y,  yrep, stat = "sd")

@@

We include one run using only variables availale in both datasets, to explore sensitivity of our results to selected covariates (Fig.\@ref(fig:kfm-data)).

```{r kfm-data, fig.cap="Model fit to PISCO data but only using variables also available for KFM data", results="hide"}

@@

We fit an alternative model using data from the KFM program to test whether a different dataset from the same region provides different results. KFM based results are qualitatively similer to those from PISCO data (Fig.\@ref(fig:kfm-fit))

```{r kfm-fit, fig.cap="Model fit to KFM data. Results are more uncertain than using PISCO data, but follow the same trend estimated from the PISCO data."}

@@

We ran a version of the model using data from only inside MPAs. This serves as a test of whether our results are being entirely driven by the outside-MPA data, and allows us to compare results from inside MPAs to those outside (Fig.\@ref(fig:mpa-fit-plot)).

@@

We ran a version of the model using data from only outside MPAs. This serves as a test of whether our results are being entirely driven by the inside-MPA data, and allows us to compare results from inside MPAs to those outside (Fig.\@ref(fig:fished-fit-plot)).

```{r fished-fit-plot, fig.cap="Estimated MPA effects using PISCO data from outside MPAs only. The base model uses data from both inside and outside MPAs. These results show that the same general trend holds when only using data from outside the MPAs, though the model estimates less probability of high positive effects witout MPA data, compared to the model fit using all the data.",results = "hide"}

@@

We include a model run using nearly all observed species, to test whether our results are being influenced by omissino of rarely observed species (Fig.\@ref(fig:no-filter-plot)).

@@

For ease of interpretation and model convergence, our default model estimates MPA effects in three-year bins. For this run we estimate the effect annually, relative to the year 2000. Results show the same overall trend and magnitude reported in our base model (Fig.\@ref(fig:no-bin-plot)).

@@

Our base run estimates the effect of MPAs on total biomass densities of targeted finfish. For this run, we instead estimate the effect on mean biomass densities of targeted species. Results show the same pattern as our base run (Fig.\@ref(fig:mean-fit)).

```{r mean-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish. For this run, we instead estimate the effect on mean biomass densities of targeted species. Results show the same pattern as our base run.", results="hide"}
mean_fit <- model_runs %>%
  slice(1) %>%
  left_join(abundance_data, by = "data_source") %>%
  mutate(
    did_fit = pmap(
      list(
        data = data,
        data_to_use = data_to_use,
        data_source = data_source
      ),
      estimate_did,
      fit_mean = TRUE,
      iter = 2500,
      site_data = site_data,
      cdfw_catches = cdfw_catches,
      life_history_data = life_history_data,
      refresh = 1000
    )
  )

@@

```{r no-zero-fit, fig.cap="Our base run estimates the effect of MPAs on total biomass densities of targeted finfish. For this run, we instead estimate the effect on mean biomass densities of targeted species. Results show the same pattern as our base run.", results = "hide"}

@@

@@

```{r betatable, results="asis"}

@@

Simulation results are presented as percent differences in biomass densities with and without MPAs, in order to be comparable to the estimates that the regression model produces. However, this metric presents some problems as a measure of how "detectable" an effect size is. As depletion increases, relatively small changes in total biomass (relative to the variance in the observation process) can translate into large percent changes in biomass. For example, moving from a density of 0.02kg/m^2^ to 0.04kg/m^2^ translates to a 100% percent increase, but only a .02kg/m^2^ absolute increase, a small value to detect with a real observation program.
